---
title: "Scaling Graphs in ggplot2"
subtitle: "how all scale_* functions work in ggplot2"
date: "2023-11-04"
format:
    html: 
      self-contained: true
      code-annotations: below
categories: [code, analysis, R]
reference-location: margin
citation-location: margin
---

## The Issue

This is an issue I faced whilst working as a developer on the dabestr rebuild sometime back. You might have also encountered this before - its the issue of "floating bar graphs" such as the one below.

```{r, warning=FALSE}
library(ggplot2)

g <- ggplot(mpg, aes(class)) +
  geom_bar()
g
```

## The Fix

```{r}
g <- ggplot(mpg, aes(class)) +
  geom_bar() +
  coord_cartesian(ylim = c(0, 60),  # <1> 
                  expand = FALSE)   # <1> 
g
```

1. Note that we used `coord_cartersian` here instead of `scale_y_continuous`. More details can be found on ggplot's [official documentation](https://ggplot2.tidyverse.org/reference/coord_cartesian.html) regarding this topic. To summarise what the documentation states, `coord_cartesian` can essentially be seen as zooming in closer to the plot whilst `scale_*` functions shrinks the limits directly, thereby eliminating possible lines or dots from being drawn.